---
title: "Chamela Adaptation Session 2"
output:
  html_document:
    df_print: paged
---

# Stochastic Linear Ornstein-Uhlenbeck models for Comparative Hypotheses

Let's suppose the predictor is not a discrete regime mapping, but a continuous predictor. Here we have salamander data
where temperatures were taken at two scales: The temperature of a salamander in its microhabitat and the "macroscale" 
temperature obtained for species via their geographic coordinates as mean average annual temperature from the BIOCLIM
data raster.

```{r}
library(slouch)
library(treeplyr)

salDat <- readRDS("../data/SalamanderMicrohabitat.rds")
salDat

```


Visualize the data with a phylomorphospace plot. 
```{r}
phytools::phylomorphospace(salDat$phy, as.matrix(salDat$dat[,c("Temp_macro_ave", "Soil_temp_ave")]), xlim=c(5, 20), ylim=c(5, 20), ftype="off")
abline(0,1, lty=2) # One to one line
```

Perform a PGLS on the data using an OU model for the residuals in phylolm.

```{r}
library(phylolm)
rownames(salDat$dat) <- salDat$phy$tip.label
pgls_fit <- phylolm(Soil_temp_ave ~ Temp_macro_ave, data=salDat$dat, phy = salDat$phy, model = "OUfixedRoot")
summary(pgls_fit)
```
Now let's fit a SLOUCH model. The first one we will fit will actually just repeat the previous exercise, and be another way of fitting a PGLS. 

```{r}
library(slouch)
slouchpgls_fit <- slouch.fit(salDat$phy, species=salDat$phy$tip.label, response = salDat[['Soil_temp_ave']], direct.cov = salDat[['Temp_macro_ave']], hillclimb = TRUE)

summary(slouchpgls_fit)
```

Next, we will allow the predictor to be a `random covariate`, or a predictor that evolves via Brownian Motion that has an effect on the primary optimum. We will 
now obtain two regressions, and "evolutionary regression" corresponding to the realized relationship, and an "optimal regression" representing what? 

```{r}
slouch_fit <- slouch.fit(salDat$phy, species=salDat$phy$tip.label, response = salDat[['Soil_temp_ave']], random.cov = salDat[['Temp_macro_ave']], hillclimb = TRUE)

summary(slouch_fit)
```
So far we have ignored measurement error. What do you predict the effect will be of measurement error on: 
1. Phylogenetic signal?
2. How will error in Y affect the slope? 
3. How will error in X affect the slope? 

```{r}
slouchme_fit <- slouch.fit(salDat$phy, species=salDat$phy$tip.label, response = salDat[['Soil_temp_ave']], mv.response = salDat[['Soil_temp_errVar']], 
                           random.cov = salDat[['Temp_macro_ave']], mv.random.cov = salDat[['Temp_macro_errVar']], hillclimb = TRUE)

summary(slouchme_fit)
```
Now let's plot the optimal and evolutionary regressions.

```{r}
plot(salDat[['Temp_macro_ave']], salDat[['Soil_temp_ave']], xlim=c(5, 20), ylim=c(5,20), xlab="Macroscale Temperature", ylab="Microhabitat Temperature")
abline(slouchme_fit$beta_primary$coefficients_bias_corr[1,1], slouchme_fit$beta_primary$coefficients_bias_corr[2,1], col="green", lwd=2)
abline(slouchme_fit$beta_evolutionary$coefficients_bias_corr[1,1], slouchme_fit$beta_evolutionary$coefficients_bias_corr[2,1], col="blue", lwd=2)
```

Let's inspect the likelihood surface. This is a key step to understanding the uncertainty in your parameters. We will show
the surface two log likelihood units below the peak, as this allows us to determine the well-supported regions of parameter space. 

```{r}
hls <- seq(0.0001, 300, length.out=40)
vys <- seq(0.01, 25, length.out=40)

slouch_gridsearch <- slouch.fit(salDat$phy, species=salDat$phy$tip.label, response = salDat[['Soil_temp_ave']], mv.response = salDat[['Soil_temp_errVar']], 
                           random.cov = salDat[['Temp_macro_ave']], mv.random.cov = salDat[['Temp_macro_errVar']], hl_values = hls, vy_values = vys)

```

```{r}
par(mar=c(1,1,1,1))
plot(slouch_gridsearch, theta=0)
plot(slouch_gridsearch, theta=-45)
plot(slouch_gridsearch, theta=45)
plot(slouch_gridsearch, phi=60)
```

That's it for now! Check out mvSLOUCH for even cooler, multivariate models. 


